---
title: "Final Project"
author: "Xi Cao"
date: "4/29/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(sf)
library(choroplethr)
library(tidyverse)
library(ggplot2)
library(magrittr)
library(dplyr)
library(tidyr)
library(readxl)
library(gutenbergr)
library(rvest)
library(tidytext)
library(ggrepel)
library(lubridate)
library(choroplethrMaps)
library(gutenbergr)
library(gridExtra)
```

Project description
In this project, we will deal with some data related to America tuition costs in each state and then find out some correlation between the tuition costs and 

Data introduction
In this project, we have used 4 data scource from Github, there is some simple description about the data we will use:
1.tuition_cost





1. Data import and data scraped
1.(a) Read data from the first data source. As data in this source has already been cleaned, so we do not have to clean them any more.
```{r read data from first source, echo=FALSE}
tuition_cost <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-03-10/tuition_cost.csv')

salary_potential <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-03-10/salary_potential.csv')

historical_tuition <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-03-10/historical_tuition.csv')

diversity_school <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-03-10/diversity_school.csv')
```

1.(b) Read data from the second source. Th data read from this source is the tuition fees and living costs by state. This data source need to do some cleaning and change some colnames. We deleted some rows and colnames and renamed the colname names. Export it as a csv file.
```{r read data from second source, echo=FALSE}
scraped <- "https://www.collegetuitioncompare.com/state/" %>%
    read_html() %>%
    html_table(fill = T) %>%  
    .[1] %>% 
    as.data.frame() 
scraped <- scraped[,-1] 
scraped <- scraped[-c(1,2),]
colnames(scraped) <- c("State", "Number of Schools", "In-State Public Tuition Fees", "Out-State Public Tuition Fees", "Private Tuition Fees", "On-Campus Living Costs", "Off-Campus Living Costs")

```

1.(c) Read more specific tuition costs from the website of each state by a loop and store these table into a list.
```{r read data from each state, echo=FALSE}
state_names <- unique(scraped$State) %>% tolower() %>% str_replace_all(" ","-")
s <- list()
for(i in 1:59){
    html <- paste0("https://www.collegetuitioncompare.com/state/", state_names[i], "/")
    s[[i]] <- html %>%
        read_html() %>%
        html_table(fill = T) %>%  
        .[1] %>% 
        as.data.frame()
}

```

1.(d) Changed the specific data collected from each state to dataframes and select data according to different tuition costs. Export these tables as csv files.
```{r change state information to a table, echo=FALSE}
library(reshape2)
s <- as.data.frame(melt(s))
s <- s[,-6]
colnames(s) <- c("Fees", "In-state Undergraduate", "Out-state Undergraduate", "In-state Graduate", "Out-state Graduate")
state_tuition <- s %>% 
    filter(Fees == "Tuition & Fees") %>% 
    mutate(State = unlist(state_names[1:59]))
state_books <- s %>% 
    filter(Fees == "Books & Supplies") %>% 
    mutate(State = unlist(state_names[1:59]))
state_room <- s %>% 
    filter(Fees == "Room & Board") %>% 
    mutate(State = unlist(state_names[1:59]))
state_total <- s %>% 
    filter(Fees == "COA") %>% 
    mutate(State = unlist(state_names[1:59]))
```



2. Diversity_school wrangling
In the part, we dealt with the diversity_school table to study the difference of enrollment between different states and categories. First, we group this table by state and category and count the total enrollment of different states and categories. Then we make a barplot and a state_choropleth to show it more specific. 
```{r diversity_school wrangling and analysis, echo=FALSE}
t1 <- diversity_school %>%
  drop_na() %>%
  group_by(state, category) %>%
  summarise(count = sum(enrollment)) %>%
  arrange(desc(count))
head(t1, n = 10)

data(df_state_demographics)

ggplot(t1) + 
    geom_bar(aes(x = count, y = state, fill = category),stat = "identity") + 
    ggtitle("Total Enrollment by State and Category")

ggplot(t1) + 
    geom_bar(aes(x = count, y = state),stat = "identity") + 
    facet_wrap(~ category) + 
    ggtitle("Total Enrollment by State and Category")

diversity_school %>% 
    drop_na() %>% 
    group_by(state) %>%
    summarise(value = sum(enrollment)) %>%
    mutate(region = tolower(state)) %>%
    inner_join(df_state_demographics, by = 'region') %>%
    state_choropleth(title = "Total Enrollment by State") + 
    theme(legend.position = "bottom")
  


```
From the table, we can see that, the category of "Total Minority" in California has the highest enrollment, and the second and third rank also come from California, that is "Women" and "Hispanic". And the fourth and fifth come from Texas which is "Women" and "Total Minority". Besides, we can learn that from the best 10 enrollment, almost half of them come from the category of "Women". 
From the first bar plot, we can see that the best 5 high enrollment states are Carlifornia, Texas, New York, Florida, Illinois and they almost have all categories which means these states have more comprehensive universities. And combined with the third plot, we can see that those states with higher total enrollment are mostly spread at the west and east coast of America where economic development is faster. 
From the second bar plot, we can see that Women, White, Total MInority, Hispanic, Black have more enrollment than other categories and Women has the highest proportion in all of these categories which is same as we have learned from the table.



3. Historical_tuition wrangling
This table show the historical tuition costs from 1985 to 2017. At this table, we only use constant dollars which is more stable with economic status. Then, we make 3 barplot to show the variation of tuition costs between different unversity types and tuition types. 
```{r historical_tuition analysis, echo=FALSE}
grid.arrange(
  historical_tuition %>% 
    mutate(type2 = str_detect(`tuition_type`, "\\bConstant")) %>%
    filter(type2 == TRUE & type == "All Institutions") %>% 
    ggplot() + 
    geom_bar(aes(x = tuition_cost, y =year , fill = tuition_type),stat = "identity", position = "dodge") + 
    ggtitle("1985-2017 All Institutions Tuition Costs"),
  historical_tuition %>% 
    mutate(type2 = str_detect(`tuition_type`, "\\bConstant")) %>%
    filter(type2 == TRUE & type == "Public") %>% 
    ggplot() + 
    geom_bar(aes(x = tuition_cost, y =year , fill = tuition_type),stat = "identity", position = "dodge") + 
    ggtitle("1985-2017 Public Tuition Costs"),
  historical_tuition %>% 
    mutate(type2 = str_detect(`tuition_type`, "\\bConstant")) %>%
    filter(type2 == TRUE & type == "Private") %>% 
    ggplot() + 
    geom_bar(aes(x = tuition_cost, y =year , fill = tuition_type),stat = "identity", position = "dodge") + 
    ggtitle("1985-2017 Private Tuition Costs"),
  ncol = 2
)
```
From the plot, although there are some missing values in time in public and private part, basically, the tuition costs has been increasing continuously from 1985 to 2017 no matter which tuition type is. Among these, private universities have the highest tuition cost, and public universities have the lowest tuition cost. Besides, all these three tuition type have increased through years, but "2 year constant" did not increased much while other two types increased over 10000 dollars through these years. 



4. salary_potential wrangling
This table shows the estimated salary according to different level of career and we will show that the difference of potential salary and university quality(make_world_better_percent: Percent of alumni who think they are making the world a better place, stem_percent: Percent of student body in STEM) between states. First, do some summarise and cleaning on the table and create two barplots and a state_choropleth which can be used to make a comparison with th eplot in step 2.
```{r salary_potential analysis, echo=FALSE}
t2 <- salary_potential %>% 
  drop_na() %>%
  group_by(state_name) %>%
  summarise(avg_early_career_pay = round(mean(early_career_pay),2),
            avg_mid_career_pay = round(mean(mid_career_pay),2),
            avg_make_world_better_perc = round(mean(make_world_better_percent),2),
            avg_stem = round(mean(stem_percent),2))

grid.arrange(
  t2 %>%
    melt() %>%
    filter(variable == "avg_early_career_pay"|variable == "avg_mid_career_pay") %>%
    ggplot() + 
    geom_bar(aes(x = value, y = state_name, fill = variable), stat = "identity", position = "dodge") +
    xlab("average career pay") +
    theme(legend.position="bottom"),
  
  t2 %>%
    melt() %>%
    filter(variable == "avg_make_world_better_perc"|variable == "avg_stem") %>%
    ggplot() + 
    geom_bar(aes(x = value, y = state_name, fill = variable), stat = "identity", position = "dodge") +
    xlab("university quality") +
    theme(legend.position="bottom"),
  ncol=2
)


t2 %>%
  rename(value = avg_mid_career_pay) %>%
  mutate(region = tolower(state_name) %>% str_replace_all("-", " ")) %>%
  inner_join(df_state_demographics, by = 'region') %>%
  state_choropleth(title = "Average Mid-Career Pay by State") + 
  theme(legend.position = "bottom")

```
First, from the two barplot, we can see that the estiamted early career pay does not differ greatly among states, it is about 40000 dollars, and the difference of make_world_better_percent between states is also not big, around 55%. But the other two variable have large difference between different states: Carlifornia, Massachusetts, New York and Pennsylvania has the highest estiamted mid-career pay, and correspondingly, these four states also have higher stem percentage compared with other states. This may mean that stem correlated majors need more professors for a university. Besides, combined the next plot with the plot "Total Enrollment by State" in part2, we also know that Carlifornia and New York have more enrollment than other states. This means enrollment and salary may have some connections, but this table only shows the estimated salary, so in the next part, we will show the relationship between enrollment and real tuition income.



5.(a) tuition_cost wrangling
This table shows the tuition cost for 2018-2019. First, we change missing values to zero. And then we analyze the tuition cost according to different classification criterion.
```{r tuition_cost analysis, echo=FALSE}
tuition_cost <- tuition_cost %>% 
  replace_na(list(state="District of Columbia", room_and_board = 0)) %>%
  mutate(total = in_state_total + out_of_state_total)

  
tuition_cost_public <- tuition_cost %>% 
  filter(type == "Public") %>%
  group_by(state, degree_length) %>%
  summarise(avg_in_state_total = round(mean(in_state_total),2),
            avg_out_of_state_total = round(mean(out_of_state_total),2),
            avg_total = round(mean(total),2)) %>%
  arrange(desc(avg_total))
head(tuition_cost_public, 5)

tuition_cost_private <- tuition_cost %>% 
  filter(type == "Private") %>%
  group_by(state, degree_length) %>%
  summarise(avg_in_state_total = round(mean(in_state_total),2),
            avg_out_of_state_total = round(mean(out_of_state_total),2),
            avg_total = round(mean(total),2)) %>%
  arrange(desc(avg_total))
head(tuition_cost_private, 5)

tuition_cost_profit <- tuition_cost %>% 
  filter(type == "For Profit") %>%
  group_by(state, degree_length) %>%
  summarise(avg_in_state_total = round(mean(in_state_total),2),
            avg_out_of_state_total = round(mean(out_of_state_total),2),
            avg_total = round(mean(total),2)) %>%
  arrange(desc(avg_total))
head(tuition_cost_profit, 5)



grid.arrange(
  ggplot(tuition_cost_private %>% melt() %>% filter(degree_length == "4 Year")) +
    geom_bar(aes(x = value, y = state, fill = variable), stat = "identity", position = "dodge") + 
    ggtitle("4 Year Tuition Cost Private") + 
    theme(legend.position="bottom"),
  ggplot(tuition_cost_private %>% melt() %>% filter(degree_length == "2 Year")) +
    geom_bar(aes(x = value, y = state, fill = variable), stat = "identity", position = "dodge") + 
    ggtitle("2 Year Tuition Cost Private") + 
    theme(legend.position="bottom"),
  ncol = 2
)

grid.arrange(
  ggplot(tuition_cost_public %>% melt() %>% filter(degree_length == "4 Year")) +
    geom_bar(aes(x = value, y = state, fill = variable), stat = "identity", position = "dodge") + 
    ggtitle("4 Year Tuition Cost Public") + 
    theme(legend.position="bottom"),
  ggplot(tuition_cost_public %>% melt() %>% filter(degree_length == "2 Year")) +
    geom_bar(aes(x = value, y = state, fill = variable), stat = "identity", position = "dodge") + 
    ggtitle("2 Year Tuition Cost Public") + 
    theme(legend.position="bottom"),
  ncol = 2
)


grid.arrange(
  ggplot(tuition_cost_profit %>% melt() %>% filter(degree_length == "4 Year")) +
    geom_bar(aes(x = value, y = state, fill = variable), stat = "identity", position = "dodge") + 
    ggtitle("4 Year Tuition Cost For Profit") +
    theme(legend.position="bottom"),
  ggplot(tuition_cost_profit %>% melt() %>% filter(degree_length == "2 Year")) +
    geom_bar(aes(x = value, y = state, fill = variable), stat = "identity", position = "dodge") + 
    ggtitle("2 Year Tuition Cost For Profit") + 
    theme(legend.position="bottom"),
  ncol = 2
)

```
The three tables show the best 5 high tuition cost according to different tuition type. we can see that Vermont always has higher tuition cost in all these three tuition types, and Rhode Island and Massachusetts is close to Vermont in private tuition type. From part4, we also know that Massachusetts has almost highest estimated potential. Besides, we can see that the tuition cost has increased compared with the histirical tuition cost in part3.
In this part, we have also created 6 bar plots to compare the difference of in-state and out-of-state under different situations which is similar to what we have done in part3. Overall, the in-state tuition cost is much small than out-of-state tuition cost. 
In private type, in-state tuition cost is the same as out-of-state tuition cost ehich is quite different from the public universities no matter 4-year or 2-year, but the 4-year tuition cost is still about twice that of 2-year. And in some states, they do not provide 2-year education. 
And compared wthin "Public" type, we can see that 4-year tuition cost is about twice that of 2-year, 4-year tuition cost is about 35000 dollars, but 2-year tuition cost is only about 12500, much smaller than 4-year.Besides, every state have public universities.
For profit type, it's the same with private type, that is, in-state tuition cost is the same as out-of-state tuition cost no matter what the degree length is. Also, there are some states don't have universities for profit type. 



5.(b) tuition_cost + diversity_school
In this part, we join these two tables to find out whether there are some coorelations between university diversity and tuition cost.
```{r tuition_cost + diversity_school, echo=FALSE}
join1 <- left_join(tuition_cost, diversity_school, by = c("name", "state")) %>%
  drop_na() %>%
  group_by(state, category) %>%
  summarise(avg_total = round(mean(total),2)) 

ggplot(join1) + 
  geom_bar(aes(x = avg_total, y = state), stat = "identity", position = "dodge") +
  facet_wrap(~ category) + 
  ggtitle("Tuition Cost on Category")
```
From the plot, we can see that there is no correlation between university diversity and tuition cost, all categories have same tuition cost in each state. But, there is some relationship between total enrollment and tuition cost, for example, Carlifornia, Texas, New York have higher enrollment and correspondingly, these states have higher tuition cost in public and profit type of tuition.



6. 2020 tuition cost data wrangling
6.(a) In this part, we will deal with the specific data we collected from each state and comibine them together. We also turn costs to numeric type for easy calculation.
```{r 2020 tuition cost data cleaning, echo=FALSE}
state_books <- state_books[,-c(1:3)]
state_books$State <- str_replace_all(state_books$State, "-", " ") 
state_books[state_books == "-"] <- 0
colnames(state_books) <- c("Undergraduate-books", "Graduate-books", "State")

state_room <- state_room[,-c(1:3)]
state_room$State <- str_replace_all(state_room$State, "-", " ") 
state_room[state_room == "-"] <- 0
colnames(state_room) <- c("Undergraduate-room", "Graduate-room", "State")

state_tuition <- state_tuition[,-1]
state_tuition$State <- str_replace_all(state_tuition$State, "-", " ") 
state_tuition[state_tuition == "-"] <- 0
colnames(state_tuition) <- c("In-state Undergraduate Tuition Fee", "Out-state Undergraduate Tuition Fee", "In-state Graduate Tuition Fee", "Out-state Graduate Tuition Fee", "State")

state_total <- state_total[,-1]
state_total$State <- str_replace_all(state_total$State, "-", " ") 
state_total[state_total == "-"] <- 0
colnames(state_total) <- c("In-state Undergraduate Total", "Out-state Undergraduate Total", "In-state Graduate Total", "Out-state Graduate Total", "State")


scraped2 <- left_join(state_total, state_tuition) %>% 
  left_join(state_room) %>% 
  left_join(state_books) %>%
  .[, c(5, 1:4, 6:13)]


scraped$`In-State Public Tuition Fees` <- as.numeric(str_replace(substr(scraped$`In-State Public Tuition Fees`, 2,10), ",",""))
scraped$`Out-State Public Tuition Fees` <- as.numeric(str_replace(substr(scraped$`Out-State Public Tuition Fees`, 2,10), ",",""))
scraped$`Private Tuition Fees` <- as.numeric(str_replace(substr(scraped$`Private Tuition Fees`, 2,10), ",",""))
scraped$`On-Campus Living Costs` <- as.numeric(str_replace(substr(scraped$`On-Campus Living Costs`, 2,10), ",",""))
scraped$`Off-Campus Living Costs` <- as.numeric(str_replace(substr(scraped$`Off-Campus Living Costs`, 2,10), ",",""))

for(i in 2:13){
  scraped2[,i] <- as.numeric(str_replace(substr(scraped2[,i], 2,10), ",",""))
  scraped2[,i] <- replace_na(scraped2[,i], 0)
}

for(i in 3:6){
  scraped[,i] <- replace_na(scraped[,i], 0)
}

write.csv(scraped, "2020 Tuition and Living Costs Summary By States")
write.csv(scraped2, "2020 Tuition and Living Costs Summary By States 2")

head(scraped, 5)
head(scraped2, 5)
```
these two tables are come from the second data source.

6.(b) Do some analysis with the data we collected from the website and find some connection with the previous analysis.
```{r data analysis with the second data source, echo=FALSE}
scraped <- scraped %>%
  .[-62,] %>%
  mutate(`Public Tuition Fees` = `In-State Public Tuition Fees` + `Out-State Public Tuition Fees`)


grid.arrange(
  scraped %>% 
    melt() %>%
    filter(variable == "public tuition fees"|variable == "Private Tuition Fees") %>%
    ggplot() + 
    geom_bar(aes(x = value, y = State, fill = variable), stat = "identity", position = "dodge") +
    ggtitle("2020 Tuition Fees By University Type") + 
    theme(legend.position="bottom"),
  
  scraped %>% 
    melt() %>%
    filter(variable == "On-Campus Living Costs"|variable == "Off-Campus Living Costs") %>%
    ggplot() + 
    geom_bar(aes(x = value, y = State, fill = variable), stat = "identity", position = "dodge") +
    ggtitle("2020 Living Costs By States") + 
    theme(legend.position="bottom"),
  ncol = 2
)

scraped2 %>% 
  select("State", "In-state Undergraduate Total", "Out-state Undergraduate Total", "In-state Graduate Total", "Out-state Graduate Total") %>%
  melt() %>%
  ggplot() + 
  geom_bar(aes(x = value, y = State), stat = "identity", position = "dodge") +
  facet_wrap(~ variable) + 
  ggtitle("2020 Costs of Attendance By States")

scraped2 %>% 
  rename(region = State, value = `Out-state Graduate Total`) %>%
  inner_join(df_state_demographics, by = 'region') %>%
  state_choropleth(title = "2020 Costs of Attendance on Graduate By States") + 
  theme(legend.position = "bottom")
```
From the first two barplot, we can know that, the cost of private type tuition is still higher than that of public type of tuition, but the difference is smaller in some states. And Vermont has the highest tuition costs both in public and private, which is similar as we have learned from part5.(a). As for living costs, there is slightly difference between on-campus living and off-campus living, Carlifornia and Hawaii have relatively higher off-campus living cost and District of Columbia, Massachusetts and Rhode Island have relatively higher on-campus living. But no matter on-campus or off-campus, it's seems like these are some relationship between tuition costs and living costs: basically, higher tuition costs may lead to higher living costs, and in these states where mostly are coastal cities, the potential salary would be higher either according to Part4. 

From the third barplot, we can see that the COA of graduate study is more higher than that of undergraduate study, and obviously, students who are not resident of this state need to pay more than those who are resident of this state. And from the last plot, we caan see that, those states with higher COA are basically coastal cities where have more international students and the economic development is faster.
